Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-01-16 paper.qmd
Updated Section 7.2 text

The MSE for different binwidths ($a_1$) using tSNE, UMAP, PHATE, PaCMAP, and TriMAP with default (hyper-)parameter setting (@fig-mnist-mse) were calculated. It is found that tSNE (@fig-mnist-mse (a)) provide the most reasonable representation for the digit 1 dataset, showing universally best. However, \gD{} representation shows a big non-linear cluster and a small cluster with a small gap (@fig-tsne-best).

In the case of digit 1 data, there should not be any clusters unless anomalies exist, indicating different digit 1 patterns. The angle of the digit 1 images varies along this non-linear clustering structure (@fig-tsne-best), while the small cluster contains the digit 1 images with different patterns of the digit 1, unlike the usual (@fig-model-error-mnist (c)). This provides the evidence for two close clusters.

By visualizing the model generated for tSNE in $10\text{-}D$ helps to assess the \gD{} layout. As shown in @fig-mnist-tri-proj (a), the model provides the evidence for the non-linear structure of the digit 1 data in $10\text{-}D$. The model shows some quirks. The model's twisted pattern provides evidence for the $10\text{-}D$ data structure, which is not observed by \gD{} layout (@fig-mnist-tri-proj (b)). Furthermore, the presence of long edges indicates the existence of the small cluster located closely together (@fig-mnist-tri-proj (c)).

>>>>

The MSE for different binwidths ($a_1$) using tSNE, UMAP, PHATE, PaCMAP, and TriMAP with default (hyper-)parameter setting (@fig-mnist-mse) were calculated. It is found that tSNE (@fig-mnist-mse (f)) provide the most reasonable representation for the digit 1 dataset, showing universally best. However, \gD{} representation shows a big non-linear cluster and a small cluster with a small gap.

In the case of digit 1 data, there should not be any clusters unless anomalies exist, indicating different digit 1 patterns. The angle of the digit 1 images varies along this non-linear clustering structure (@fig-model-error-mnist), while the small cluster contains the digit 1 images with different patterns of the digit 1, unlike the usual (@fig-model-error-mnist (c)). This provides the evidence for two close clusters.

By visualizing the model generated for tSNE in $10\text{-}D$ helps to assess the \gD{} layout. As shown in @fig-mnist-tri-proj (a1), the model provides the evidence for the non-linear structure of the digit 1 data in $10\text{-}D$. The model shows some quirks. The model's twisted pattern provides evidence for the $10\text{-}D$ data structure, which is not observed by \gD{} layout (@fig-mnist-tri-proj (a2)).
